\bfseries A Variant of the Sandwich theorem for two-sided limits

### 2025-07-01 lecturenotes_aima203/ch6_aima203.tex
changes

\subsection{Newton's forward difference formulae}
The Newton's forward difference formula is:
\begin{equation}
  \label{newtforward6}
 \begin{gathered}
  y_n(x) = y_0 + u \Delta y_0 + \frac{u(u-1)}{2!}\, \Delta ^2 y_0 + \frac{u(u-1)(u-2)}{3!}\, \Delta ^3 y_0 + ... \\[1mm]
  + \frac{u(u-1)(u-2)...(u-(n-1))}{n!}\, \Delta ^n y_0
\end{gathered}
\end{equation}
where, $\mathbf{x=x_0 + uh}$ and $h = x_{i+1} - x_{i}$. Differentiating ~\ref{newtforward6} with respect to $x$,

\begin{equation}
  \label{newtforwarddif}
  \begin{aligned}
    \frac{dy}{dx} =\frac{dy}{du} \frac{du}{dx} = \frac{1}{h} \left [ \Delta y_0 + \frac{2u -1}{2}\, \Delta^2\,  y_0 + \frac{3u^2-6u + 2}{6}\, \Delta^3 \, y_0 + ... \right ]
  \end{aligned}
\end{equation}
Differentiating ~\ref{newtforwarddif} with respect to $x$, we get,
\begin{equation}
  \label{newtforwarddif2}
  \begin{aligned}
    \frac{d^2y}{dx^2} = \frac{1}{h^2} \left [ \Delta^2 y_0 + \frac{6u -6}{6}\, \Delta^3 \,  y_0 + \frac{12u^2-36u + 22}{24}\, \Delta^4 \, y_0 + ... \right ]
  \end{aligned}
\end{equation}
These formulae are used for \textit{non-tabular} values of $x$. For tabular values of $x$, the formulae take simpler form. For $x=x_0$, we have $u=0$, and using this we can find the relations of $\displaystyle \left [ \frac{dy}{dx} \right ]  _{x=x_0} and \left [\frac{d^2y}{dx^2} \right ] _{x=x_0}$ which is in a simpler form.
\newpage
\begin{equation}
  \label{newtforw1x}
  \left [ \frac{dy}{dx} \right ] _{x=x_n} = \frac{1}{h} \left [\Delta \, y_n - \frac{1}{2} \Delta ^2 \, y_n + \frac{1}{3} \Delta ^3 \, y_n - ... \right ]
\end{equation}
and
\begin{equation}
  \label{newtforw2x}
  \left [ \frac{d^2y}{dx^2} \right ] _{x=x_n} = \frac{1}{h^2} \left [\Delta^2 \, y_n - \Delta ^3 \, y_n + \frac{11}{12} \Delta ^4 \, y_n - \frac{5}{6} \Delta ^5 \, y_n + ... \right ]
\end{equation}

>>>>

\subsection{Newton's forward difference formulae}
The Newton's forward difference formula is:
\begin{equation}
  \label{newtforward6}
 \begin{gathered}
  y_n(x) = y_0 + p \Delta y_0 + \frac{p(p-1)}{2!}\, \Delta ^2 y_0 + \frac{p(p-1)(p-2)}{3!}\, \Delta ^3 y_0 + ... \\[1mm]
  + \frac{p(p-1)(p-2)...(p-(n-1))}{n!}\, \Delta ^n y_0
\end{gathered}
\end{equation}
where, $\mathbf{x=x_0 + ph}$ and $h = x_{i+1} - x_{i}$. Differentiating ~\ref{newtforward6} with respect to $x$,

\begin{equation}
  \label{newtforwarddif}
  \begin{aligned}
    \frac{dy}{dx} =\frac{dy}{dp} \frac{dp}{dx} = \frac{1}{h} \left [ \Delta y_0 + \frac{2p -1}{2}\, \Delta^2\,  y_0 + \frac{3p^2-6p + 2}{6}\, \Delta^3 \, y_0 + ... \right ]
  \end{aligned}
\end{equation}
Differentiating ~\ref{newtforwarddif} with respect to $x$, we get,
\begin{equation}
  \label{newtforwarddif2}
  \begin{aligned}
    \frac{d^2y}{dx^2} = \frac{1}{h^2} \left [ \Delta^2 y_0 + \frac{6p -6}{6}\, \Delta^3 \,  y_0 + \frac{12p^2-36p + 22}{24}\, \Delta^4 \, y_0 + ... \right ]
  \end{aligned}
\end{equation}
These formulae are used for \textit{non-tabular} values of $x$. For tabular values of $x$, the formulae take simpler form. For $x=x_0$, we have $p=0$, and using this we can find the relations of $\displaystyle \left [ \frac{dy}{dx} \right ]  _{x=x_0} and \left [\frac{d^2y}{dx^2} \right ] _{x=x_0}$ which is in a simpler form.
\newpage
\begin{equation}
  \label{newtforw1x}
  \left [ \frac{dy}{dx} \right ] _{x=x_n} = \frac{1}{h} \left [\Delta \, y_n - \frac{1}{2} \Delta ^2 \, y_n + \frac{1}{3} \Delta ^3 \, y_n - ... \right ]
\end{equation}
and
\begin{equation}
  \label{newtforw2x}
  \left [ \frac{d^2y}{dx^2} \right ] _{x=x_n} = \frac{1}{h^2} \left [\Delta^2 \, y_n - \Delta ^3 \, y_n + \frac{11}{12} \Delta ^4 \, y_n - \frac{5}{6} \Delta ^5 \, y_n + ... \right ]
\end{equation}

### 2025-07-01 sandwich_theorem/extended_sandwich_theorem.tex
changes

\vspace{5mm}
\noindent
\textbf{Abstract:} {The criterion for the classical Sandwich theorem for two-sided limits is that the two-side limits must exists for the bounding functions. We show that this criterion can be relaxed. We show that it is sufficient for the existence of left-hand limit for lower bound function and the existence of right-hand limit for the upper bound function; and of-course they must be equal. This paper relaxes the criterion of the classical Sandwich theorem by replacing the two-sided limits with one-sided limits in the criterion and gives a new variant of the Sandwich theorem.  While Rudin has formulated the Sandwich theorem for the one-sided limits, it still doesn't relax the criterion of the Sandwich theorem for two-sided limits. He has incorporated the one-sided limits for the Sandwich theorem for one-sided limits but has not relax the condition for the Sandwich theorem for two-sided limits.}
\section{Introduction}
The Sandwich theorem is simple yet powerful tool in analysis to determine and to analyze the limit of a function at a given point. We can leverage the known limits to calculate the unknown limits. Suppose, we know the limits of \(g(x)\) and \(h(x)\) at \(x=c\) to be the same limit \(L\) and here \(f(x)\) happens to be sandwich between \(g(x)\) and \(h(x)\) in some neighborhood of \(c\). Then we can conclude the limit of \(f(x)\) at \(x=c\) as \(L\) by the Sandwich theorem for two-sided limits. The Sandwich theorem for two-sided limits is simply called the Sandwich theorem which is as follows.

@@

 \section{Remarks on the our variant of Sandwich theorem}
 \begin{itemize}
 \item  While the \(\displaystyle \lim_{x \to 0} \frac{sinx}{x}\) can be determined by the sandwich theorem using the tighter-bound $cosx \leq sinx/x \leq 1$, the purpose of our variant of sandwich theorem is not to assert that this limit cannot be determined by the Sandwich theorem. This limit can be determined without the sandwich theorem as well.

>>>>

\vspace{5mm}
\noindent
\textbf{Abstract:} {The criterion for the classical Sandwich theorem for two-sided limits is that the two-side limits must exists for the bounding functions. We show that this criterion can be relaxed. We show that it is sufficient for the existence of the left-hand limit for the lower bound function and the existence of right-hand limit for the upper bound function; and of-course they must be equal. This paper relaxes the criterion of the classical Sandwich theorem by replacing the two-sided limits with one-sided limits in the criterion and gives a new variant of the Sandwich theorem. While Rudin has formulated the Sandwich theorem for the one-sided limits, it still doesn't relax the criterion of the Sandwich theorem for two-sided limits. He has incorporated the one-sided limits for the Sandwich theorem for one-sided limits but has not relax the condition for the Sandwich theorem for two-sided limits~\cite{rudin}.}
\section{Introduction}
The Sandwich theorem is simple yet powerful tool in analysis to determine and to analyze the limit of a function at a given point. We can leverage the known limits to calculate the unknown limits. Suppose, we know the limits of \(g(x)\) and \(h(x)\) at \(x=c\) to be the same limit \(L\) and here \(f(x)\) happens to be sandwich between \(g(x)\) and \(h(x)\) in some neighborhood of \(c\). Then we can conclude the limit of \(f(x)\) at \(x=c\) as \(L\) by the Sandwich theorem for two-sided limits. The Sandwich theorem for two-sided limits is simply called the Sandwich theorem which is as follows.

@@

 \section{Remarks on the our variant of Sandwich theorem}
 \begin{itemize}
 \item  While the \(\displaystyle \lim_{x \to 0} \frac{sinx}{x}\) can be determined by the sandwich theorem using the tighter-bound $cosx \leq sinx/x \leq 1$ ~\cite{tao}, the purpose of our variant of sandwich theorem is not to assert that this limit cannot be determined by the Sandwich theorem. This limit can be determined without the sandwich theorem as well.
